---
title: "Define the number of PCs to use using a function"
author: "Martin Loza
date: "13/04/2021"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=5, 
  fig.height=5
)
```

We first load the BicPCA and other packages

```{r setup}
library(BicPCA)
library(ggplot2)
library(Seurat)
library(here)

source(here("Functions.r"))

seed = 777
```

We load a datasat from tabula muris

```{r}
load(file = here("Data/Tabula_Muris/facs_Thymus_seurat_tiss.Robj"))
x <- UpdateSeuratObject(tiss)
rm(tiss)
x
```
# Standard way to define the number of PCs

According with the elbow plot, we would like to use around 20 dimensions, but this is a really subjective way to do it. Minka (2000) provides an alternative way using the Bayesian Information Criterion (BIC) over a probabilistic PCA. We then need to calculate the BIC for a desired set of PCs and define the number of PCs as the one with the minimum BIC score.

```{r}
x <- NormalizeData(x, verbose = FALSE)
x <- FindVariableFeatures(x, verbose = FALSE)
x <- ScaleData(x, verbose = FALSE)
x <- RunPCA(x, npcs = 70, verbose = FALSE)
ElbowPlot(x, ndims = 70)
x
```
# Define the number of PCs using the BIC

We now define the number of PCs to use. If we set returnBIC = TRUE, we get all the BIC for each tested dimension. Plotting the BIC could be useful for debbuging.    

```{r} 
bicPCA <- FindNumberPC(x, dim2Test = 100, returnBIC = TRUE)
selPCA <- bicPCA$selPCA
plot(bicPCA$BIC[1:30])
print(paste0("Selected PCs: ", selPCA))
```
We can use the selected number of PCs in downstream analysis as visualization. 

```{r}
x <- RunUMAP(x, reduction = "pca", dims = 1:selPCA, verbose = FALSE)
```

```{r}
DimPlot(x, group.by = "cell_ontology_class", label = TRUE, label.size = 5) + NoLegend()
ElbowPlot(x)
```












