---
title: "Seurat-objects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Seurat-objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We first load the BicPCA package

```{r setup}
library(BicPCA)
library(ggplot2)
library(Seurat)
library(here)

seed = 777
```

We load stored data from tabula muris

```{r}

load(file = here("Data/Tabula_Muris/facs_Thymus_seurat_tiss.Robj"))
x <- UpdateSeuratObject(tiss)
rm(tiss)
x
```

# Standard workflow

According with the elbow plot, we would like to use around 20 dimensions, but this is a really subjective way to do it. Minka (2000) provides an alternative way using the Bayesian Information Criterion (BIC) over a probabilistic PCA. We then need to calculate the BIC for a desired set of PCs and define the number of PCs as the one with the minimum BIC score.

```{r}
x <- NormalizeData(x, verbose = FALSE)
x <- FindVariableFeatures(x, verbose = FALSE)
x <- ScaleData(x, verbose = FALSE)
x <- RunPCA(x, npcs = 70, verbose = FALSE)
ElbowPlot(x, ndims = 70)
x
```

To use the BIC to define the number of PCs we need the standard deviations of the complete PCA's dimensions (e.g., if we use 2000 variable features, we need the 2000 standad deviation of the corresponding PCA).  This is a problem for Seurat objects as it uses the package irlba to compute the PCA and we just btain the s.d. of the selected PCs  

```{r}
length(x[["pca"]]@stdev)
```

The good thing is that in order to use the BIC we can provide the eigen values of the covariance matrix of the centered data. The eigen values are comparable on 3 digits. Good enough for our tests

```{r}
scaledData <- x@assays$RNA@scale.data
eig1 <- (prcomp(x = t(scaledData), center = FALSE, scale. = FALSE)$sdev)^2
eig2 <- eigen(cov(t(scaledData)))$values

sum(eig1 == eig2[1:length(eig1)])
sum((eig1 - eig2[1:length(eig1)]) < 1e-3)
```

Now we use the getBic function with the eigen values to calculate the BIC. According with this criterion, we will select the number of PCs that corresponds with the minimum BIC value. We can observe that there is a difference if we use the eigen values obtained from the functio prcomp repect with the ones obtained from the covariance matrix. This might be related to the fact that using the prcomp function we obtained only 1349 components, given by the fact that we have larger dimensions than samples. However, both selection of PCs agree, even though they are in a different scale. The results agree with the Elbow Plot.

```{r}
BIC1 <- -BicPCA::getBicPca(eig = eig1, N = ncol(scaledData), K = 20)
BIC2 <- -BicPCA::getBicPca(eig = eig2, N = ncol(scaledData), K = 20)
plot(BIC1)
plot(BIC2)
ElbowPlot(x)
selPCA <- which.min(BIC1)
```

Finally we can use the selected number of PCs in downstream analysis as visualization. 

```{r}
x <- RunUMAP(x, reduction = "pca", dims = 1:selPCA, verbose = FALSE)
```

```{r}
DimPlot(x, group.by = "cell_ontology_class")
ElbowPlot(x)
```












