---
title: "Seurat-objects"
author: "Martin Loza"
---

On this workflow I will explore dotplots.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=5, 
  fig.height=5
)
```

Load some libraries

```{r setup }
library(ggplot2)
library(Seurat)
library(here)
library(tidyverse)
library(cowplot)

seed = 777

# Scales the plot color gradient to a given scale (limits) and colors.
ScaleGradient <- function(plot, colLow = "gray", colHigh = "purple", limits = NULL){
  if(is.null(limits)){
    stop("No defined limits. Select the limits and try again.", call. = TRUE)
  }
  plot <- plot + scale_color_gradient(low = colLow, high = colHigh, limits = limits)  
  
  return(plot)
}
```

We load the FACS data from [tabula muris](https://ndownloader.figshare.com/files/13092395) 

```{r}
load(file = here("Data/Tabula_Muris/facs_Thymus_seurat_tiss.Robj"))
x <- UpdateSeuratObject(tiss)
x <- NormalizeData(x, verbose = FALSE)
x <- FindVariableFeatures(x, verbose = FALSE)
rm(tiss)
x
```

# Arrange the data

Lets arrange the data to plot normalized counts for now. We need:
 * *avg_count* the gene average expression per group (e.g. celltype, cluster, etc.)
 * *n_cells_group* the total number of cells from each group
 * *n_expressing_cells* the number of cells with detectable (>0) expression for each gene per cluster.
Let's start with the gene average expression per group. To ease the use of different data in the future, we will associate the data with a general label "avg_count". We need the group labels to split the data, find the average and join the results back into a single data frame.

```{r}
# sample to ease processing
idxCells <- sample(1:ncol(x), size = 300, replace = FALSE)
selGenes <- VariableFeatures(x)[1:10] # use the first 10 genes

datal <- SplitObject(x[selGenes, idxCells], split.by = "cell_ontology_class")
namesl <- names(datal)
df <- lapply(namesl, function(n){
  
  tmp <- t(as.matrix(GetAssayData(datal[[n]], assay = "RNA", slot = "data")))
  df <- as.data.frame(colMeans(tmp))
  colnames(df) <- "avg_count"
  
  expressing <- colSums(tmp > 0)
  
  df[["n_expressing_cells"]] <- expressing[rownames(df)]
  df[["n_cells_group"]] <- nrow(tmp)
  df <-  mutate(df, per_expressing = (n_expressing_cells/n_cells_group)*100)
  df[["group"]] <- n
  df[["gene"]] <- rownames(df)
  
  order <- c("gene", "avg_count", "group", "per_expressing", "n_cells_group", "n_expressing_cells")
  df <- df[order]
  
  return(df)
})
df <- df %>% reduce(rbind)
df
```

## Dotplot 

We can create the dotplot and compare with the seurat ones. The difference from Seurat is that they "un-normalize" the data using the `expm1` function, then find the average expression and then normalize again using the `log1p` function. In my case, I just take the average of the already normalized counts, (used by scanpy... I think). It would be interesting to check how this change affects. Anyways, because mainly we are interested on visualize differences, both options are acceptable.
 
```{r fig.width=7}
df$gene <- factor(df$gene, levels = selGenes )
p <- ggplot(df, aes(x=group, y = gene, color = avg_count, size = per_expressing)) + 
  geom_point() + 
  coord_flip() + 
  theme_cowplot() + 
  labs(col = "Average expression", size = "Percentage expressing") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p <- ScaleGradient(p, colHigh = "blue", limits = c(0, max(df$avg_count))) 
p

DotPlot(x[,idxCells], group.by = "cell_ontology_class",  features = selGenes, scale = FALSE) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
We can create functions to use in further analyses.

```{r}
GetDfDotPlot <- function(object = NULL, genes = NULL, group = NULL, type = "avg_logcounts", scale = FALSE, ...){
  
  if(is.null(genes)){
    warning("No genes provided. The top 10 variable features will be used.", call. = TRUE)
    features <- Seurat::VariableFeatures(object)[1:10]
    
    if(is.null(features)){
      object <- Seurat::FindVariableFeatures(object, verbose = FALSE)
      features <- Seurat::VariableFeatures(object)[1:10]
    }
  }else{
    features <- genes
  }
  
  object <- object[features,]
  
  if(!is.null(group)){
    datal <- Seurat::SplitObject(object, split.by = group)
  }else{
    datal = list(object)
    names(datal) <- "object"
  }
  
  if(type == "avg_scaled")
    scale = TRUE
  
  namesl <- names(datal)
  df <- lapply(namesl, function(n){
    
    if(type == "avg_counts"){
      tmp <- t(as.matrix(Seurat::GetAssayData(datal[[n]], assay = "RNA", slot = "counts")))
      avg_expression <- colMeans(tmp)
    }else if(type == "avg_logcounts" || type == "avg_scaled"){
      tmp <- t(as.matrix(Seurat::GetAssayData(datal[[n]], assay = "RNA", slot = "data")))
      avg_expression <- colMeans(tmp)
    }else if(type == "seurat"){
      tmp <- t(as.matrix(Seurat::GetAssayData(datal[[n]], assay = "RNA", slot = "data")))
      avg_expression <- colMeans(expm1(tmp))
      avg_expression <- log1p(avg_expression)
    }else if(type == "avg_scaled"){
      tmp <- t(as.matrix(Seurat::GetAssayData(datal[[n]], assay = "RNA", slot = "scale.data")))
      avg_expression <- colMeans(tmp)
    }
    
    expressing <- colSums(tmp > 0)
    
    df <- as.data.frame(avg_expression)
    
    df[["n_expressing_cells"]] <- expressing[rownames(df)]
    df[["n_cells_group"]] <- nrow(tmp)
    df <-  plyr::mutate(df, per_expressing = (n_expressing_cells/n_cells_group)*100)
    df[["group"]] <- n
    df[["gene"]] <- rownames(df)
    
    order <- c("gene", "avg_expression", "group", "per_expressing", "n_cells_group", "n_expressing_cells")
    df <- df[order]
    
    return(df)
  })
  df <- df %>% reduce(rbind)
  
  if(scale == TRUE){
    df$avg_expression <- scale(df$avg_expression)
  }
  
  df$gene <- factor(df$gene, levels = features )
  
  return(df)
}

GetDotPlot <- function(object = NULL, genes = NULL, group = NULL, scale = FALSE, type = "avg_logcounts", ...){
  if(is.null(object))
    stop("Error. Object need to be defined.", call. = TRUE)
  
  df <- GetDfDotPlot(object = object, genes = genes, group = group, type = type, scale = scale)
  
  p <- ggplot(df, aes(x=group, y = gene, color = avg_expression, size = per_expressing)) + 
    geom_point() + 
    coord_flip() + 
    theme_cowplot() + 
    labs(col = "Average expression", size = "Percentage expressing") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  p <- ScaleGradient(p, limits = c(min(df$avg_expression), max(df$avg_expression)), ...) 
  return(p)
}
```

Using the functions.

```{r}
GetDotPlot(object = x, genes = selGenes, group = "cell_ontology_class", colHigh = "steelblue" )
```
CONCLUSIONS: Seurat's DotPlot is useful and easy to use, the problem it's that we dont' have much flexibility to change the information displayes. By using my own function, I am free to use any info I need by just changing some short code lines. 






