---
title: "Filter pairs"
author: "Martin Loza"
---

On this workflow I will filter the outlier pairs using the IQR measure.

## Init

```{r setup}
library(Canek)
library(CanekDebug)
library(Seurat)
library(here)
library(RNAseqAnalysis)
library(patchwork)
library(ggplot2)

seed = 777

# Functions
histPairsDistance <- function(object = NULL, cluster = NULL, plotMean = FALSE, plotMedian = FALSE, colorMean = "red", colorMedian = "blue", ...){
  
  pairs <- CanekDebug::GetMNNPairsData(object)
  cluster <- CanekDebug::GetMembershipData(object)[[cluster]]
  
  # Subset the pairs
  idxCells <- which(pairs$query %in% cluster$cells)
  pairs <- pairs[idxCells,]
  
  data <- as.matrix(Seurat::GetAssayData(object, assay = "RNA", slot = "data"))
  
  # Get the pairs distance
  distance <- sapply(seq_len(nrow(pairs)), function(idx){
    queCell <- pairs[idx,"query"]
    refCell <- pairs[idx,"ref"]
    
    queCell <- data[, queCell, drop = FALSE]
    refCell <- data[, refCell, drop = FALSE]
    
    pair <- cbind(queCell, refCell)
    return(dist(t(pair)))
  })
  
  df <- data.frame(distance = distance)
  p <- ggplot(df, aes(x = distance,...)) + geom_histogram(...) + xlab("Pairs distance")
  
  if(plotMean)
    p <- p +  geom_vline(xintercept = mean(df$distance), color = colorMean)
  if(plotMedian)
    p <- p + geom_vline(xintercept = median(df$distance), color = colorMedian)
  
  return(p)
}
```

## Load data

```{r}
xl <- Canek::SimBatches$batches
nameBatches <- names(xl)
xl <- lapply(nameBatches, function(n){
  counts <- xl[[n]]
  object <- Seurat::CreateSeuratObject(counts = counts, project = "CanekDebug")
  object[["batch"]] <- n
  return(object)
})

x <- Reduce(merge, xl)

x[["celltype"]] <- Canek::SimBatches$cell_types
table(x[["batch"]])
table(x[["celltype"]])
```

## Pre-processing

```{r}
set.seed(seed)
x <- RNAseqAnalysis::SeuratPreprocessing(x)
```

## Umap before correction

The batch-effect causes the cell-type-1 cells to group by batch

```{r}
set.seed(seed)
x <- RNAseqAnalysis::GetUMAP(object = x, dims = 10)
DimPlot(x, group.by = "celltype") + DimPlot(x, group.by = "batch")
```

## Correction with Canek

Canek returns the correction information if `debug = TRUE`.

```{r}
set.seed(seed)
x <- Canek::RunCanek(x = x, batches = "batch", debug = TRUE, kNN = 500)
```

## UMAP after correction

```{r}
set.seed(seed)
x <- RNAseqAnalysis::GetUMAP(object = x, dims = 10, assay = "Canek")
DimPlot(x, group.by = "celltype") + DimPlot(x, group.by = "batch")
```

## Correction info

For this test we will analyse the MNN pairs from cluster 1.

```{r}
pairs <- CanekDebug::GetMNNPairsData(x)
cluster <- CanekDebug::GetMembershipData(x)[[1]]

# Subset the pairs
idxCells <- which(pairs$query %in% cluster$cells)
pairs <- pairs[idxCells,]
```

We get the normalized data before correction, and calculate the cells pairs distance.

```{r}
data <- as.matrix(Seurat::GetAssayData(x, assay = "RNA", slot = "data"))

distance <- sapply(seq_len(nrow(pairs)), function(idx){
  
  queCell <- pairs[idx,"query"]
  refCell <- pairs[idx,"ref"]
  
  queCell <- data[, queCell, drop = FALSE]
  refCell <- data[, refCell, drop = FALSE]
  
  pair <- cbind(queCell, refCell)
  return(dist(t(pair)))
})
```

## Quantiles 

The median corresponds to the 50% quantile.

```{r}
q <- quantile(distance)
q
mean(distance)
median(distance)
```

The interquantile range is given by Q1(25%) - 1.5(IQR), Q3(75%) + 1.5*IQR, where IQR = Q3-Q1.

```{r}
IQR <- q["75%"] - q["25%"]
lowOutliers <- which(distance < (q["25%"] - (1.5*IQR)))
highOutliers <- which(distance > (q["75%"] + (1.5*IQR)))
outliers <- c(lowOutliers, highOutliers)
```

## Distance plots

```{r fig.width=10, fig.height=5}
binwidth = 0.05
df1 <- data.frame(distance = distance)
df2 <- data.frame(distance = distance[-outliers])

p1 <- ggplot(df1, aes(x = distance)) +
  geom_histogram(binwidth =  binwidth, alpha = 0.8) +
  xlab("Pairs distance") + geom_vline(xintercept = mean(df1$distance), color = "red") + 
  geom_vline(xintercept = median(df1$distance), color = "blue") + ggtitle(label = "Without filtering", subtitle = paste0("Mean: ", mean(df1$distance), ", Median: ", median(df1$distance)))

p2 <- ggplot(df2, aes(x = distance)) +
  geom_histogram(binwidth = binwidth, alpha = 0.8) +
  xlab("Pairs distance") + geom_vline(xintercept = mean(df2$distance), color = "red") + 
  geom_vline(xintercept = median(df2$distance), color = "blue") + ggtitle(label = "With filtering", subtitle = paste0("Mean: ", mean(df2$distance), ", Median: ", median(df2$distance)))

p1 + p2
```






## Session info
```{r}
sessionInfo()
```




