---
title: "Using scMerge with Seurat objects"
author: "Martin Loza"
---

In this test I will use scMerge with Seurat objects and compare different parameters selection. 

```{r setup}
library(here)
library(Seurat)
library(scMerge)
library(ggplot2)

options(future.globals.maxSize = 4e10)

seed = 321

# Functions
SeuratPreprocessing <- function(object = NULL,...){
  object <- NormalizeData(object, ...)
  object <- FindVariableFeatures(object, ...)
  
  return(object)
}

GetPCA <- function(object = NULL, npcs = 10, ...){
  object <- ScaleData(object, ...)
  object <- RunPCA(object, npcs = npcs, ...)
  
  return(object)
}
```

## Load data and Create Seurat Object

For this test I will use the Jurkat/293t/50:50 datasets from 10x genomics. 

```{r}
Jurkat <- readRDS(here("Data/Jurkat/Jurkat.Rds"))
t293 <- readRDS(here("Data/Jurkat/t293.Rds"))
Mix50_50 <- readRDS(here("Data/Jurkat/Mix50_50.Rds"))

xl <- list(Jurkat = Jurkat,
           t293 = t293,
           Mix50_50 = Mix50_50)

x <- Reduce(merge, xl)
x
table(x$batch)
```

## Preprocessing

Normalization and find variable features.

```{r}
x <- SeuratPreprocessing(x, verbose = FALSE)
```

## PCA before correction

Batch effects cause the Jurkat cells to group by batch. 

```{r}
x <- GetPCA(x, npcs = 30, verbose = FALSE)
DimPlot(x, group.by = "celltype")
ElbowPlot(x)
dimPCA <- 5
```

## Using scMerge with Seurat objects

scMerge works with bioconductor, so I need to convert the Seurat object. I used the found variable features to standardize the genes of the assays.

```{r}
features <- VariableFeatures(x)
sceObject <- Seurat::as.SingleCellExperiment(x[features,])
sceObject
```

Also I need to define negative controls, which I think they could be the variable features from the seurat object. I would like to compare the correction using the variable features and the negative controls found by the function `scSEGIndex`. For this coparison test I will use the k's for each batch (e.g. the number of known clusters per batch) as 1, just for comparison.

```{r}
data <- SummarizedExperiment::assay(sceObject, 'logcounts')
seg = scMerge::scSEGIndex(exprs_mat = data)
k = c(1,1,1)
group <- "batch"
assayName <- "scMerge"

tFea <- scMerge::scMerge(sce_combine = sceObject, ctl = features, 
                            assay_name = assayName, kmeansK = k,
                            batch_name = group, plot_igraph = FALSE,
                            verbose = FALSE)

tSEG <- scMerge::scMerge(sce_combine = sceObject, ctl = rownames(seg), 
                            assay_name = assayName, kmeansK = k,
                            batch_name = group, plot_igraph = FALSE,
                            verbose = FALSE)

xFea <- CreateSeuratObject(counts = as.matrix(SummarizedExperiment::assay(tFea, "scMerge")),
                           assay = "integrated",
                           meta.data = x[[]])
xSEG <- CreateSeuratObject(counts = as.matrix(SummarizedExperiment::assay(tSEG, "scMerge")),
                           assay = "integrated",
                           meta.data = x[[]])

VariableFeatures(xFea) <- features
VariableFeatures(xSEG) <- rownames(seg)

xFea <- GetPCA(object = xFea, npcs = 10, assay ="integrated", verbose = FALSE)
xSEG <- GetPCA(object = xSEG, npcs = 10, assay ="integrated", verbose = FALSE)

xFea <- RunUMAP(xFea, dims = 1:10, verbose = FALSE)
xSEG <- RunUMAP(xSEG, dims = 1:10, verbose = FALSE)
```

We can observe that, even though the correction is not succesful, probably because of the ks selected, the selection of negative controls doesn't look to have a high impact on the results.

```{r}
DimPlot(xFea, group.by = "batch")
DimPlot(xSEG, group.by = "batch")
```

Now I will perform the same test using k as [1,1,2], which corresponds to the number of known celltypes on each dataset.

```{r}
k = c(1,1,2)
group <- "batch"
assayName <- "scMerge"

tFea <- scMerge::scMerge(sce_combine = sceObject, ctl = features, 
                            assay_name = assayName, kmeansK = k,
                            batch_name = group, plot_igraph = FALSE,
                            verbose = FALSE)

tSEG <- scMerge::scMerge(sce_combine = sceObject, ctl = rownames(seg), 
                            assay_name = assayName, kmeansK = k,
                            batch_name = group, plot_igraph = FALSE,
                            verbose = FALSE)

xFea <- CreateSeuratObject(counts = as.matrix(SummarizedExperiment::assay(tFea, "scMerge")),
                           assay = "integrated",
                           meta.data = x[[]])
xSEG <- CreateSeuratObject(counts = as.matrix(SummarizedExperiment::assay(tSEG, "scMerge")),
                           assay = "integrated",
                           meta.data = x[[]])

VariableFeatures(xFea) <- features
VariableFeatures(xSEG) <- rownames(seg)

xFea <- GetPCA(object = xFea, npcs = 10, assay ="integrated", verbose = FALSE)
xSEG <- GetPCA(object = xSEG, npcs = 10, assay ="integrated", verbose = FALSE)

xFea <- RunUMAP(xFea, dims = 1:10, verbose = FALSE)
xSEG <- RunUMAP(xSEG, dims = 1:10, verbose = FALSE)
```

The corrections look good now. Maybe the correction using the SEG the clusters looks tighter. I would prefer the correction using the variable features, but it is just my personal and subjective opinion.

```{r}
DimPlot(xFea, group.by = "batch") + ggtitle("Features")
DimPlot(xSEG, group.by = "batch") + ggtitle("SEG")
```

## Wrapper function

```{r}
RunScMerge <- function(seuratObject = NULL, group = "batch", assayName = "integrated", k = NULL, runningTime = FALSE){
  
  md <- seuratObject[[]]
  features <- VariableFeatures(seuratObject)
  
  if(length(features) == 0)
    stop('No variable features found. Run Seurat::FindVariableFeatures and try again.')
  
  object <- Seurat::as.SingleCellExperiment(seuratObject[features,])
  
  if(is.null(k)){
    k <- rep(5, length(unique(md[[batch]])))
  }
  
  time <- system.time({
    object <- scMerge::scMerge(sce_combine = object,
                               ctl = features,
                               assay_name = assayName,
                               kmeansK = k,
                               batch_name = group,
                               plot_igraph = FALSE,
                               verbose = FALSE)
  })
  
  #Create Seurat object
  object <- CreateSeuratObject(counts = as.matrix(SummarizedExperiment::assay(object, assayName)),
                               assay = assayName,
                               meta.data = seuratObject[[]])
  
  VariableFeatures(object) <- features
  
  if(runningTime == FALSE)
    return(object)
  else
    return(list(scMerge = object, time = time))
}
```

Now we can use the wrapper function directly with a Seurat object.

```{r}
x <- RunScMerge(seuratObject = x, group = "batch", k = c(1,1,2))
x <- GetPCA(x, npcs = dimPCA, assay ="integrated", verbose = FALSE)
x <- RunUMAP(x, dims = 1:dimPCA, verbose = FALSE)
```

```{r}
DimPlot(x, group.by = "batch")
```
CONCLUSIONS:
The effect of using SGE or features does not have a high impact on the results. However, we MUST select the appropiate k's for the internal k-means when using scMerge. This is the main drawback of this method. The results are highly affected by the selection of these k's, which in most cases are un-known.

## Session info

```{r}
sessionInfo()
```



