---
title: "Test scale data"
author: "Martin Loza
date: "13/04/2021"
---

GOAL: I want to manually obtain the same results as compare with Seurat's scaling if I do it manually

We first load the BicPCA and other packages

```{r setup}
library(BicPCA)
library(ggplot2)
library(Seurat)
library(here)

source(here("Functions.r"))

seed = 777
set.seed(seed)
```

We load a datasat from tabula muris

```{r}
load(file = here("Data/Tabula_Muris/facs_Thymus_seurat_tiss.Robj"))
x <- UpdateSeuratObject(tiss)
rm(tiss)
x
```
We first calculate the typical workflow with Seurat objects

```{r}
x <- NormalizeData(x, verbose = FALSE)
x <- FindVariableFeatures(x, verbose = FALSE)
x <- ScaleData(x, verbose = FALSE)

scaDataSeu <- x@assays$RNA@scale.data
```

Now we scale the data ussing the normalized data over the variable features

```{r}
features <- VariableFeatures(x)
scaDataMan <- as.matrix(GetAssayData(x, assay = "RNA", slot = "data")[features,])

scaDataMan <- t(scale(t(scaDataMan), center = TRUE, scale = TRUE))
```

We compare the two scalings. We first re-check that we are using the same variable features on both scalings. After comparing, we don't obtained the sames values...

```{r}
scaDataMan <- scaDataMan[rownames(scaDataSeu),]
sum(rownames(scaDataMan) %in% rownames(scaDataSeu)) == length(features)

expected <- length(scaDataSeu)

sum(scaDataMan == scaDataSeu)
sum(scaDataMan == scaDataSeu) == expected
sum(abs(scaDataMan - scaDataSeu) < 1e-1)
sum(abs(scaDataMan - scaDataSeu) < 1e-1) == expected
```
```{r}
scaDataMan <- as.matrix(GetAssayData(x, assay = "RNA", slot = "data")[rownames(scaDataSeu),])
scaDataMan <- t(scale(t(scaDataMan), center = TRUE, scale = TRUE))
```

CONCLUSIONS: We need to search more about the way that Seurat is scaling the data...











