---
title: "Wrapper for ComBat"
author: "Martin Loza"
---

On this workflow I will test the ComBat method to correct batch-effects. My final goal is to create a wrapper function to use with Seurat objects.

## Init

```{r, setup}
library(here)
library(sva)
library(Seurat)
library(ggplot2)

dimPCA <- 10
seed = 777
sample = TRUE

source(here("Functions.R"))
```

## Load data and create Seurat Object

For this test I will use the Jurkat/293t/50:50 datasets from 10x genomics. I sampled the datasets to ease the experiments.

```{r}
Jurkat <- readRDS(here("Data/Jurkat/Jurkat.Rds"))
t293 <- readRDS(here("Data/Jurkat/t293.Rds"))
Mix50_50 <- readRDS(here("Data/Jurkat/Mix50_50.Rds"))

xl <- list(Jurkat = Jurkat,
           t293 = t293,
           Mix50_50 = Mix50_50)

x <- Reduce(merge, xl)

sampleData <- function(object, frac = NULL, seed = 777, ...){
  set.seed(seed)
  nCells <- ncol(object)
  samples <- floor(frac*nCells)
  idx <- sample(x = nCells, size = samples, ... )
  
  return(object[,idx])
}

if(sample)
  x <- sampleData(x, frac = 0.2, seed = seed)
  
x
table(x$batch)
```

## Preprocessing

Normalization and find variable features.

```{r}
set.seed(seed)
x <- SeuratPreprocessing(x, verbose = FALSE)
```

## PCA before correction

Batch effects cause the Jurkat cells to group by batch. 

```{r}
set.seed(seed)
x <- GetPCA(x, nPCA = 30)
DimPlot(x, group.by = "celltype")
ElbowPlot(x)
```
## ComBat

To use ComBat we need to pass the normalized count matrix (gene x sample) and the batch labels. 

```{r}
object <- x
features <- Seurat::VariableFeatures(object)
data <- as.matrix(Seurat::GetAssayData(object, assay = "RNA", slot = "data")[features,])
md <- object[[]]

corrData <- sva::ComBat(dat = data, batch = factor(md[["batch"]]))

object[["integrated"]] <- Seurat::CreateAssayObject(counts = corrData)
DefaultAssay(object) <- "integrated"
VariableFeatures(object) <- features
object
```
```{r}
object <- GetPCA(object = object, nPCA = dimPCA)
DimPlot(object, group.by = "celltype") + ggtitle("Corrected")
```
## Wrapper function

```{r}
RunComBat <- function(object = NULL, batch = "batch", runningTime = FALSE, verbose = FALSE, ...){

  features <- Seurat::VariableFeatures(object)
  if(length(features) == 0){
    warning("Variable features not defined. Running 'FindVariableFeatures' function.", call. = TRUE)
    features <- Seurat::VariableFeatures(Seurat::FindVariableFeatures(object, verbose = verbose))
  }

  data <- as.matrix(Seurat::GetAssayData(object, assay = "RNA", slot = "data")[features,])
  md <- object[[]]

  if(!(batch %in% colnames(md)))
    stop(paste0(batch, "not found in object's metadata. Check the batch label."))

  time <- system.time({
    corrData <- sva::ComBat(dat = data, batch = md[[batch]], ...)
  })

  object[["integrated"]] <- Seurat::CreateAssayObject(counts = corrData)
  Seurat::DefaultAssay(object) <- "integrated"
  Seurat::VariableFeatures(object) <- features

  if(runningTime == FALSE)
    return(object)
  else
    return(list(object = object, time = time))
}
```

Testing the wrapper function.

```{r}
x <- RunComBat(object = x, batch = "batch")
x <- GetPCA(object = x, nPCA = dimPCA)
DimPlot(x, group.by = "celltype") + ggtitle("Normalized")
```
CONCLUSIONS: ComBat requires a good definition of the variable features to use, different features could lead to an erroneous correction.  

## Session info
```{r}
sessionInfo()
```








