---
title: "Wrapper for ComBat-seq"
author: "Martin Loza"
---

On this workflow I will test the ComBat-seq method to correct batch-effects. My final goal is to create a wrapper function to use with Seurat objects.

## Init

```{r, setup}
library(here)
library(sva)
library(Seurat)
library(ggplot2)

dimPCA <- 5
seed = 777
sample = TRUE

source(here("Functions.R"))
```

## Load data and create Seurat Object

For this test I will use the Jurkat/293t/50:50 datasets from 10x genomics. I sampled the datasets to ease the experiments.

```{r}
Jurkat <- readRDS(here("Data/Jurkat/Jurkat.Rds"))
t293 <- readRDS(here("Data/Jurkat/t293.Rds"))
Mix50_50 <- readRDS(here("Data/Jurkat/Mix50_50.Rds"))

xl <- list(Jurkat = Jurkat,
           t293 = t293,
           Mix50_50 = Mix50_50)

x <- Reduce(merge, xl)

sampleData <- function(object, frac = NULL, seed = 777, ...){
  set.seed(seed)
  nCells <- ncol(object)
  samples <- floor(frac*nCells)
  idx <- sample(x = nCells, size = samples, ... )
  
  return(object[,idx])
}

if(sample)
  x <- sampleData(x, frac = 0.2, seed = seed)
  
x
table(x$batch)
```

## Preprocessing

Normalization and find variable features.

```{r}
set.seed(seed)
x <- SeuratPreprocessing(x, verbose = FALSE)
```

## PCA before correction

Batch effects cause the Jurkat cells to group by batch. 

```{r}
set.seed(seed)
x <- GetPCA(x, nPCA = 30)
DimPlot(x, group.by = "celltype")
ElbowPlot(x)
```
## ComBat-seq

To use ComBat-seq we need to pass the raw count matrix (gene x sample) and the batch labels.

```{r}
counts <- as.matrix(GetAssayData(x, assay = "RNA", slot = "counts"))
md <- x[[]]

corrCounts <- sva::ComBat_seq(counts = counts, batch = md$batch, full_mod = FALSE)

x[["integrated"]] <- Seurat::CreateAssayObject(counts = corrCounts)
DefaultAssay(x) <- "integrated"
x
```

According to the description of ComBat-seq, the corrected matrix would be a matrix of corrected "counts", so we would need to normalize them. To test this feature I plotted the PCA of the normalized and non-normalized data. As shown in the plots, the plot without normalization is quite different from the original uncorrected PCA plot. The plot using normalized data looks similar to the original plot. These results indicate that we *need* to normalize the corrected matrix. 

```{r}
norm.x <- SeuratPreprocessing(object = x, assay = "integrated")
norm.x <- GetPCA(object = norm.x, nPCA = dimPCA)

notnorm.x <- FindVariableFeatures(object = x, assay = "integrated", verbose = FALSE)
notnorm.x <- GetPCA(object = notnorm.x, nPCA = dimPCA)

DimPlot(norm.x, group.by = "celltype") + ggtitle("Normalized")
DimPlot(notnorm.x, group.by = "celltype") + ggtitle("Non-normalized")
```
## Wrapper function

```{r}
RunComBatseq <- function(object = NULL, batch = "batch", runningTime = FALSE, verbose = FALSE, ...){
  
  features <- VariableFeatures(object)
  if(length(features) == 0){
    warning("Variable features not defined. Running 'FindVariableFeatures' function.", call. = TRUE)
    features <- VariableFeatures(FindVariableFeatures(object, verbose = verbose))
  }
    
  counts <- as.matrix(GetAssayData(object, assay = "RNA", slot = "counts")[features,])
  md <- object[[]]
  
  if(!(batch %in% colnames(md)))
    stop(paste0(batch, "not found in object's metadata. Check the batch label."))
  
  time <- system.time({
    corrCounts <- sva::ComBat_seq(counts = counts, batch = md[[batch]], full_mod = FALSE)
  })
  
  object[["integrated"]] <- Seurat::CreateAssayObject(counts = corrCounts)
  DefaultAssay(object) <- "integrated"
  
  object <- NormalizeData(object = object, assay = "integrated", verbose = verbose, ...)
  VariableFeatures(object) <- features
  
  if(runningTime == FALSE)
    return(object)
  else
    return(list(object = object, time = time))
}
```

Testing the wrapper function.

```{r}
x <- RunComBatseq(object = x, batch = "batch")
x <- GetPCA(object = x, nPCA = dimPCA)
DimPlot(x, group.by = "celltype") + ggtitle("Normalized")
```
CONCLUSIONS: ComBat-seq methods input and output data are counts. It is important to normalize the output data for its usage in further steps. 

## Session info
```{r}
sessionInfo()
```








