---
title: "Wrapper for Harmony"
author: "Martin Loza"
---

On this workflow I will test the Harmony method to correct batch-effects. My final goal is to create a wrapper function to use with Seurat objects.

## Init

```{r, setup}
library(here)
library(harmony)
library(Seurat)
library(ggplot2)

dimPCA <- 10
seed = 777
sample = TRUE

source(here("Functions.R"))
```

## Load data and create Seurat Object

For this test I will use the Jurkat/293t/50:50 datasets from 10x genomics. I sampled the datasets to ease the experiments.

```{r}
Jurkat <- readRDS(here("Data/Jurkat/Jurkat.Rds"))
t293 <- readRDS(here("Data/Jurkat/t293.Rds"))
Mix50_50 <- readRDS(here("Data/Jurkat/Mix50_50.Rds"))

xl <- list(Jurkat = Jurkat,
           t293 = t293,
           Mix50_50 = Mix50_50)

x <- Reduce(merge, xl)

sampleData <- function(object, frac = NULL, seed = 777, ...){
  set.seed(seed)
  nCells <- ncol(object)
  samples <- floor(frac*nCells)
  idx <- sample(x = nCells, size = samples, ... )
  
  return(object[,idx])
}

if(sample)
  x <- sampleData(x, frac = 0.2, seed = seed)
  
x
table(x$batch)
```

## Preprocessing

Normalization and find variable features.

```{r}
set.seed(seed)
x <- SeuratPreprocessing(x, verbose = FALSE)
```

## PCA before correction

Batch effects cause the Jurkat cells to group by batch. 

```{r}
set.seed(seed)
x <- GetPCA(x, nPCA = 30)
DimPlot(x, group.by = "celltype")
ElbowPlot(x)
```
## Harmony

We can directly pass a Seurat object to `RunHarmony` function, with the requirement that PCA must be calculated in advanced. 

```{r}
object <- x
object <- RunPCA(object = object, npcs = dimPCA, assay = "RNA", verbose = FALSE)
object <- harmony::RunHarmony(object = object, group.by.vars = "batch", dims.use	
= 1:dimPCA, verbose = FALSE)
DimPlot(object, group.by = "celltype", reduction = "harmony") + ggtitle("Corrected")
```
## Wrapper function

```{r}
RunHarmony <- function(object = NULL, batch = "batch", dims = 10, runningTime = FALSE, verbose = FALSE, ...){

  if(!("pca" %in% Seurat::Reductions(object))){
    if(verbose)
      print("Running PCA.")
    time <- system.time({
      object <- RNAseqAnalysis::GetPCA(object = object, dims = dims, verbose = verbose, ...)
      object <- harmony::RunHarmony(object = object, group.by.vars = batch, dims.use = 1:dims, verbose = verbose, ...)
    })
  }else{
    time <- system.time({
      object <- harmony::RunHarmony(object = object, group.by.vars = batch, dims.use = 1:dims, verbose = verbose, ...)
    })
  }
    
  if(runningTime == FALSE)
    return(object)
  else
    return(list(object = object, time = time))
}
```

Testing the wrapper function.

```{r}
x <- RunHarmony(object = object, batch = "batch", dims = dimPCA)
DimPlot(x, group.by = "celltype", reduction = "harmony") + ggtitle("Normalized")
```
CONCLUSIONS: Harmony is easy to use with Seurat objects. Its results are nice! :)   

## Session info
```{r}
sessionInfo()
```








