---
title: "Wrapper for Liger"
author: "Martin Loza"
---

On this workflow I will test the Liger method to correct batch-effects. My final goal is to create a wrapper function to use with Seurat objects.

## Init

```{r, setup}
library(here)
library(rliger)
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(RNAseqAnalysis)

dimPCA <- 5
seed = 777
sample = TRUE
```

## Load data and create Seurat Object

For this test I will use the Jurkat/293t/50:50 datasets from 10x genomics. I sampled the datasets to ease the experiment.

```{r}
Jurkat <- readRDS(here("Data/Jurkat/Jurkat.Rds"))
t293 <- readRDS(here("Data/Jurkat/t293.Rds"))
Mix50_50 <- readRDS(here("Data/Jurkat/Mix50_50.Rds"))

xl <- list(Jurkat = Jurkat,
           t293 = t293,
           Mix50_50 = Mix50_50)

x <- Reduce(merge, xl)

if(sample)
  x <- RNAseqAnalysis::SampleData(object = x, frac = 0.2, seed = seed)
  
x
table(x$batch)
```

## Preprocessing

Normalization and find variable features.

```{r}
set.seed(seed)
x <- SeuratPreprocessing(x, verbose = FALSE)
```

## PCA before correction

Batch effects cause the Jurkat cells to group by batch. 

```{r}
set.seed(seed)
x <- GetPCA(object = x, dims = 30)
DimPlot(x, group.by = "celltype")
ElbowPlot(x)
```
## Liger on Seurat objects 

`SeuratWrappers` already implements functions to use Liger on Seurat objects. I will make use of these functions to create my wrapper. [Liger](http://htmlpreview.github.io/?https://github.com/welch-lab/liger/blob/master/vignettes/Integrating_multi_scRNA_data.html) scales the data but do not center it. They also recommend to define the number of factor `k` in the range of 20-40, and the regularization parameter `lambda` as the default value 5. I will use `k = 30` (e.g. the mean value in the recommended range) and the default `lambda = 5`.

```{r}
set.seed(seed)
object <- x[VariableFeatures(x),]
object <- Seurat::ScaleData(object, split.by = "batch", do.center = FALSE, verbose = FALSE)
object <- SeuratWrappers::RunOptimizeALS(object,  k = 30, lambda = 5, split.by = "batch")
object <- SeuratWrappers::RunQuantileNorm(object, split.by = "batch")
```

Liger returns a low-dimensional representation of the data named `iNMF`. These reduction can be use in downstream analyses. To ease interpretation I will rename these reduction to `Liger`

```{r}
set.seed(seed)
object[["Liger"]] <- object[["iNMF"]]
dimLiger <- ncol(object[["Liger"]])
object <- GetUMAP(object = object, reduction = "Liger", dims = dimLiger, PCA = FALSE, scale = FALSE)
DimPlot(object, group.by = "celltype")
```
## Wrapper function

```{r}
RunLiger <- function(object = NULL, batch = "batch", k = 30, runningTime = FALSE, verbose = FALSE, ...){
  
  features <- Seurat::VariableFeatures(object)
  
  if(length(features) == 0){
    warning("Variable features not defined. Running 'FindVariableFeatures' function.", call. = TRUE)
    features <- Seurat::VariableFeatures(Seurat::FindVariableFeatures(object, verbose = verbose))
  }
  
  if(!(batch %in% colnames(object[[]])))
    stop(paste0(batch, "not found in object's metadata. Check the batch labels."))
  
  tmp <- object[features,]
  
  time <- system.time({
    tmp <- Seurat::ScaleData(tmp, split.by = "batch", do.center = FALSE, verbose = verbose, ...)
    tmp <- SeuratWrappers::RunOptimizeALS(tmp, k = k, split.by = "batch", ...)
    tmp <- SeuratWrappers::RunQuantileNorm(tmp, split.by = "batch", ...)
  })
  
  object[["Liger"]] <- tmp[["iNMF"]]
  
  if(runningTime == FALSE)
    return(object)
  else
    return(list(object = object, time = time))
}
```

Testing the wrapper function.

```{r}
set.seed(seed)
x <- RunLiger(object = x, batch = "batch")
```

```{r}
set.seed(seed)
dimLiger <- ncol(x[["Liger"]])
x <- GetUMAP(object = x, reduction = "Liger", dims = dimLiger, PCA = FALSE, scale = FALSE)
DimPlot(x, group.by = "celltype")
```

CONCLUSIONS: Even though we need to perform three steps, the Seurat's wrapper functions considerable eased the implementation of Liger on Seurat objects. 

## Session info

```{r}
sessionInfo()
```








