---
title: "Batch correcion using Canek with Bioconductor"
author: "Martin Loza"
date: "03/06/2021"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.align = "center",
  collapse = TRUE,
  comment = "#>"
)
```

On this test I will implement Canek using Bioconductor objects

```{r setup}
library(Canek)
library(SingleCellExperiment)
library(scran)
library(scater)
library(BiocSingular)
library(irlba)

theme_set(theme_classic())

# Functions
getPCA <- function(sce.object = NULL, n = 3, assay = "logcounts"){
  pca <- assays(sce.object)[[assay]]
  pca <- prcomp_irlba(t(pca), n = n ,center = TRUE, scale. = TRUE)
  reducedDims(sce.object) <- list(PCA = pca$x)
  return(sce.object)
}

plotPCA <- function(x, tag = "batch"){
  tag <- x[[tag]]
  pca <- data.frame(reducedDim(x, "PCA"))
  pca$tag <- tag
  
  ggplot(pca, aes(x = PC1, y = PC2, color = tag)) + geom_point()
}
```

# Load data

```{r}
simData <- Canek::SimBatches
counts <- cbind(simData$batches$B1, simData$batches$B2)
batch <- c(rep("B1", ncol(simData$batches$B1)), rep("B2", ncol(simData$batches$B2)))
x <- SingleCellExperiment(assays = list(counts = counts),
                          colData = data.frame(batch = batch, celltype = simData$cell_types))

x
```

```{r}
table(x$batch)
```

# Pre-processing

## Normalization

```{r}
x <- computeSumFactors(x)
x <- logNormCounts(x)
x
```

## Variance modelling

```{r}
dec <- modelGeneVar(x)
top <- getTopHVGs(dec, n=100)
```

### Visualizing the fit

```{r}
# fit <- metadata(dec)
# plot(fit$mean, fit$var, xlab="Mean of log-expression",
#     ylab="Variance of log-expression")
# curve(fit$trend(x), col="dodgerblue", add=TRUE, lwd=2)
```

# PCA

```{r}
x <- getPCA(x, n = 3)
x
```

```{r}
plotPCA(x, tag = "batch") + ggtitle("Batch")
plotPCA(x, tag = "celltype") + ggtitle("Celltype")
```

# Batch effect correction with Canek

```{r}
x <- RunCanek(x, batches = "orig.ident",  assay = "logcounts", features = features)
```

# PCA

```{r}
x <- getPCA(x, n = 3,  assay = "Canek")
x
```

```{r}
plotPCA(x, tag = "orig.ident") + ggtitle("Batch")
plotPCA(x, tag = "cell_type") + ggtitle("Celltype")
```
CONCLUSIONS:
The integration is unsuccessful, there is a big difference between Canek's correction using a Seurat or a SingleCellExperiment object. I think is related with the selected features to use in the integration, but I need to recheck and debug. 







