---
title: "Mutual Nearest Neighbors"
author: "Martin Loza"
---

On this workflow I will implement the MNN algorithm. The goal is to create a function to find the MNN from a pairs of matrices

## Init

```{r}
library(Canek)
library(FNN)
seed = 777 # Lucky seed
```

## Load data

I will use the dataset from Canek package.

```{r}
datal <- Canek::SimBatches$batches

sizeM1 <- 7
set.seed(seed)
idx <- sample(ncol(datal[[1]]), size = sizeM1, replace = FALSE)
datal[[1]] <- datal[[1]][1:2,idx]

sizeM2 <- 3
set.seed(seed)
idx <- sample(ncol(datal[[2]]), size = sizeM2, replace = FALSE)
datal[[2]] <- datal[[2]][1:2,idx]

m1 <- t(datal[[1]])
rownames(m1) <- 1:sizeM1
m2 <- t(datal[[2]])
rownames(m2) <- 1:sizeM2

m1
m2
```

The FNN package provides a function to calculate the nearest neighbors from two matrices. The function returns a matrix with the NN indexes and their distances. The columns of the indexes matrix corresponds to the indexes of the `data` matrix, and the rows corresponds with the indexes of the `query` matrix.
I will make a test using 2 nearest neighbors. 

```{r}
k = 2
NN <- FNN::get.knnx(data = m2, query = m1, k = k)
NN$nn.index
NN$nn.dist
nrow(NN$nn.index) == nrow(m1)
```

It's easier if we create a data frame.

```{r}
nObservations <- nrow(m1)
df <- data.frame(matrix(data = integer(), ncol = k, nrow = nObservations*k))
colnames(df) <- c("m1", "m2")
df[["m1"]] <- rep(x = seq_len(nObservations), each = k)
df[["m2"]] <- Reduce(rbind, t(NN$nn.index))
df[["distance"]] <- Reduce(rbind, t(NN$nn.dist))
NN$nn.index
NN$nn.dist
df
```

Let's check the pairs using the distances. They match!!

```{r}
distances <- as.matrix(dist(rbind(m1,m2)))
distances[1:7,8:10]

d <- sqrt((m1[1,1]- m2[1,1])^2 + (m1[1,2]- m2[1,2])^2)
trunc(d,4) == trunc(distances[1,1],4)
d <- sqrt((m1[5,1]- m2[3,1])^2 + (m1[5,2]- m2[3,2])^2)
trunc(d,4) == trunc(distances[5,3],4)
```

## Function

```{r}
FindkNN <- function(m1 = NULL, m2 = NULL,k = NULL){
  nObservations <- nrow(m1)
  df <- data.frame(matrix(data = integer(), ncol = (k+1), nrow = nObservations*k))
  colnames(df) <- c("m1", "m2", "distance")

  NN <- FNN::get.knnx(data = m2, query = m1, k = k)
  df[["m1"]] <- rep(x = seq_len(nObservations), each = k)
  df[["m2"]] <- Reduce(rbind, t(NN$nn.index))
  df[["distance"]] <- Reduce(rbind, t(NN$nn.dist))
  
  return(df)
}
```

Using the function

```{r}
df2 <- FindkNN(m1 = m1, m2 = m2, k = 2)
identical(df, df2)
df
df2
```
## MNN

Now I have to find the crossed NN.

```{r}
m1m2NN <- FindkNN(m1 = m1, m2 = m2, k = 2)
m2m1NN <- FindkNN(m1 = m2, m2 = m1, k = 2)
colnames(m2m1NN) <- c("m2","m1", "distance")

m1m2NN
m2m1NN
length(unique(m1m2NN$m1)) == nrow(m1)
length(unique(m2m1NN$m1)) == nrow(m2)
```

To find the MNN we need to check whether a kNN pair is repeated. It is convinient then to bind the dfs.

```{r}
m2m1NN <- m2m1NN[,colnames(m1m2NN)]

df <- rbind(m1m2NN, m2m1NN)
df
colnames(m1m2NN) == colnames(m2m1NN)
nrow(df) == (sizeM1 + sizeM2)*k
```

Now we create a label to distinguish repeated pairs.

```{r}
df[["pair"]] <- paste(df$m1, df$m2, sep = "_")

# short test
mnns1 <- df$pair[duplicated(df$pair)]
mnns2 <- names(which(table(df$pair) > 1))
mnns2 %in% mnns1
mnns1
```

Filtering non MNN pairs

```{r}
dfMNN <- df[duplicated(df$pair),]
rownames(dfMNN) <- dfMNN$pair
dfMNN <- dfMNN[,c("m1", "m2", "distance")]
dfMNN
rownames(dfMNN) %in% mnns1
```

## Function

```{r}
FindMNN <- function(m1 = NULL, m2 = NULL, k = NULL){
  
  ## Obtain the crossed kNN pairs
  m1m2NN <- FindkNN(m1 = m1, m2 = m2, k = k)
  m2m1NN <- FindkNN(m1 = m2, m2 = m1, k = k)
  ## Accomodate 
  colnames(m2m1NN) <- c("m2","m1", "distance")
  m2m1NN <- m2m1NN[,colnames(m1m2NN)]
  df <- rbind(m1m2NN, m2m1NN)
  ## Find MNN
  df[["pair"]] <- paste(df$m1, df$m2, sep = "_")
  df <- df[duplicated(df$pair),]
  rownames(df) <- df$pair
  df <- df[,c("m1", "m2", "distance")]
  
  return(df)
}
```

Testing the function

```{r}
df <- FindMNN(m1 = m1, m2 = m2, k = 2)
df
df$m1 %in% dfMNN$m1
identical(df, dfMNN)
```

## Test with bigger data

```{r}
datal <- Canek::SimBatches$batches

m1 <- t(datal[[1]])
m2 <- t(datal[[2]])

m1[1:6,1:6]
m2[1:6,1:6]
```

```{r}
df <- FindMNN(m1 = m1, m2 = m2, k = 10)
head(df)
nrow(df)
```



## Session info

```{r}
sessionInfo()
```







