---
title: "Scaling with matrixStats package"
author: "Martin Loza"
---

Test for scaling using the matrixStats package

## Init 

```{r}
library(matrixStats)
```

## Load data

```{r}
data <- iris[-5]
```

According with https://www.r-bloggers.com/2016/02/a-faster-scale-function/ the scaling by rows is faster.

We first center the data 
 
```{r}
data_transformed <- t(data) #Accomodate by rows
rm <- rowMeans(x = data_transformed, na.rm = TRUE)
head(rm)
```

Now we center by substracting the means. The row means will be zero (or close)

```{r}
data_transformed <- data_transformed - rm
head(rowMeans(data_transformed))
```

Now we scale the rows. We divide the rows by their standard deviations and the 

```{r}
rsd <- rowSds(data_transformed, na.rm = TRUE)
data_transformed <- data_transformed/rsd
head(rowSds(data_transformed))
```

What happens if one row is just zeros?
Centering is not affected.

```{r}
data_transformed <- t(data)
data_transformed[1,] <- 0
rm <- rowMeans(x = data_transformed, na.rm = TRUE)
data_transformed <- data_transformed - rm
head(rowMeans(data_transformed))
```

The problem is scaling.... because the sd is 0, then when we divide it produces NAs. We would need to remove those zero rows.

```{r}
rsd <- rowSds(data_transformed, na.rm = TRUE)
head(rsd)
data_transformed <- data_transformed/rsd
head(rowSds(data_transformed))
```

## Filtering 

```{r}
# Setup the test
data_transformed <- t(data)
data_transformed[1,] <- 0

## Init
cNames <- colnames(data_transformed)
rNames <- rownames(data_transformed)

## Get only zero rows 
idx <- which(matrixStats::rowSums2(data_transformed) == 0)
zeroName <- rNames[idx]
zeroName

zeroIdx <- !(rNames %in% zeroName)
data_transformed <- data_transformed[zeroIdx,, drop = FALSE ]
colnames(data_transformed) <- cNames

rm <- rowMeans(x = data_transformed, na.rm = TRUE)
data_transformed <- data_transformed - rm
head(rowMeans(data_transformed))
rsd <- rowSds(data_transformed, na.rm = TRUE)
head(rsd)
data_transformed <- data_transformed/rsd
head(rowSds(data_transformed))
```

## Functions

```{r}
# Find rows with only zeros.
WhichRowZeros <- function(matrix = NULL, idxReturn = FALSE){
  
  ## Init
  rNames <- rownames(matrix)
  results <- list()
  
  if(length(rNames) == 0)
    rNames <- seq(1:nrow(matrix))
  
  ## Get row zeros names 
  idx <- which(matrixStats::rowSums2(matrix) == 0)
  rowZeros <- rNames[idx]
  
  results[["names"]] <- rowZeros
  
  if(idxReturn)
    results[["idx"]] <- idx
  
  return(results)
}

# Filter rows with zeros
FilterRowZeros <- function(matrix = NULL, rowZeros = NULL){
  
  ## Init
  rNames <- rownames(matrix)
  
  if(length(rNames) == 0)
    rNames <- seq(1:nrow(matrix))
  
  if(is.null(rowZeros))
    rowZeros <- WhichRowZeros(matrix = matrix, idxReturn = TRUE)$idx
  
  matrix <- matrix[-zeroIdx,, drop = FALSE]
  
  return(matrix)
}

# Scale by rows
RowScale <- function(matrix, center = TRUE, scale = FALSE, filterRowZeros = FALSE){
  
  if((filterRowZeros == FALSE) & scale){
    if(length(WhichRowZeros(matrix = matrix)$names) != 0)
      stop("Rows with only zeros were found. Scaling is not possible.")
  }
  
  if(filterRowZeros){
    matrix <- FilterRowZeros(matrix = matrix)
  }
  
  if(center){
    rm <- matrixStats::rowMeans2(x = matrix, na.rm = TRUE)
    matrix <- matrix - rm
  }
  
  if(scale){
    rsd <- matrixStats::rowSds(matrix, na.rm = TRUE)
    matrix <- matrix/rsd
  }
  
  return(matrix)
}
```

Test the functions

```{r}
# Setup the test
data_transformed <- t(data)
data_transformed[1,] <- 0

WhichRowZeros(matrix = data_transformed, idxReturn = FALSE)
WhichRowZeros(matrix = data_transformed, idxReturn = TRUE)
```

```{r}
rowZeros <- WhichRowZeros(matrix = data_transformed, idxReturn = FALSE)

t1 <- FilterRowZeros(matrix = data_transformed, rowZeros = rowZeros)
t2 <- FilterRowZeros(matrix = data_transformed)

identical(t1, t2)
```

```{r}
sds <- c(1,1,1)
rm <- rowMeans2(data_transformed[-1,] - rowMeans2(data_transformed[-1,]))
#RowScale(matrix = data_transformed, center = TRUE, scale = TRUE, filterRowZeros = FALSE)

t <- RowScale(matrix = data_transformed, center = TRUE, scale = TRUE, filterRowZeros = TRUE)
nrow(t) == 3
(rowSds(t) - sds) < 1e-5
(rowMeans2(t) - rm) < 1e-5

#RowScale(matrix = data_transformed, center = TRUE, scale = TRUE, filterRowZeros = FALSE)
t <- RowScale(matrix = data_transformed, center = TRUE, scale = FALSE, filterRowZeros = TRUE)
nrow(t) == 3
(rowMeans2(t) - rm) < 1e-5

t <- RowScale(matrix = data_transformed, center = TRUE, scale = FALSE, filterRowZeros = FALSE)
nrow(t) == 4
rowSds(t) != c(1,1,1,1)
rowMeans2(t) != c(0,0,0,0)

t <- RowScale(matrix = data_transformed, center = FALSE, scale = TRUE, filterRowZeros = TRUE)
nrow(t) == 3
rowSds(t) != c(1,1,1)
rowMeans2(t) != c(0,0,0)

t <- RowScale(matrix = data_transformed, center = FALSE, scale = TRUE, filterRowZeros = TRUE)
nrow(t) == 3
rowSds(t) != c(1,1,1)
rowMeans2(t) != c(0,0,0)

#t <- RowScale(matrix = data_transformed, center = FALSE, scale = TRUE, filterRowZeros = FALSE)

t <- RowScale(matrix = data_transformed, center = FALSE, scale = FALSE, filterRowZeros = FALSE)
identical(t, data_transformed)
```
## Session info

```{r}
sessionInfo()
```











